SMT-Based Generation of Symbolic Automata \thanks{This work was partially funded by the Associated Team FM4CPS between INRIA and ECNU, Shanghai}

### 2020-01-14 2017-Xudong-SMT/ActaInformatica.tex
Acta Inf paper: minor revisions following latest reviews

Figure~\ref{schema:ArchFailure:BIP} shows this modified version of the
Failure Monitor architecture used in~\cite{CubETH-case-study}.\footnote{%
%
  \Simon{%
  The original CubETH case study \cite{CubETH-case-study} focused on
  the possibility of assembling a model of a complex software system
  in a systematic way by applying architectures to discharge
  individual system requirements.  Here, we are more interested in the
  properties of the Failure Monitor architecture by itself.  For that
  reason, we provide a modified version, in particular decomposing the
  coordinator into a Controller and a Timer, altering them to allow
  more flexibility in the acceptable behaviours.  In other words, this
  modification enforces essentially the same properties, while
  discarding less acceptable behaviours.
  }
%
} The architecture consists of 2 \emph{coordinator components}
($\NameCtrl$ and $\NameTimer$, for \emph{Control} and \emph{Timer},
respectively), 5 \emph{dangling ports} ($\PortFail$, $\PortResume$,
$\PortTick$, $\PortReset$ and $\PortAsk$) and 5 connectors.
%
%% Contrary to standard BIP models, architectures comprise one or several
%% \emph{operand} components, whereof only the set of \emph{ports} is
%% given.  Here, the operand component is $\NameOprnd$ and its interface
%% consists of the ports $\PortResume$ and $\PortFail$.
%% The two \emph{coordinator} components\mdash $\NameCtrl$ and
%% $\NameTimer$\mdash are standard
The  \emph{behaviour} of BIP components %% insofar as they also
%% have their 
is specified by finite automata extended with
local data variables.  Transitions of these automata are labelled with
the ports of the corresponding components, Boolean guards and update
functions on local variables.  %% For instance, the loop transition
%% $t_1 \goesto{\PortTick, [t < Max], t:=t+1} t_1$
%% in the $\NameTimer$ component is labelled by the port $\PortTick$. It
%% can be fired only when the current value of the local variable $t$ is
%% lower than $Max$.  Upon firing, this transition increments the value
%% of $t$ by $1$.  When omitted, the default guard (\resp update
%% function) is the constant predicate $\true$ (\resp the $\mathit{skip}$
%% operator).  The constant $\mathrm{Max}$ above, is a parameter of the
%% architecture.
Although this is not visible in the figure, below we will assume that
the dangling ports $\PortFail$ and $\PortResume$ belong to the operand
component, whereas the dangling ports $\PortTick$, $\PortAsk$ and
$\PortReset$ represent actions of the environment.

@@

We validate the approach on \Simon{a larger example}, %\st{two larger examples}, 
based on a BIP architecture taken from an earlier nanosatellite case
study~\cite{CubETH-case-study}.  \Simon{This example shows that
open-automata-based semantics can be instrumental in verifying the
properties enforced by the architectures through an encoding into open
pNets.
In order to evaluate the potential
%
of our approach to scale to larger systems, we also apply the proposed
techniques to a BIP architecture obtained by combining two instances
of the Failure Monitor architecture.  Our results show that, even
though, the performance of the smart algorithm degrades more than
linearly, execution times remain reasonably small.  It should be noted
that the aim of our approach, particularly in the case of BIP
architectures, is to allow symbolic verification of design patterns to
show that they do impose corresponding properties. Safety of the system is obtained compositionally, whereas liveness is verified by checking \emph{pair-wise} non-interference. In both cases, the main characteristic of the approach is compsitionality, so we can compute the behavior and analyse the properties on \emph{small open systems}, and use these properties independently when composing bigger systems.
%
}
%% \st{We also show how such architectures
%% can be composed, and how our approach scales.
%% This example shows that
%% open-automata-based semantics can be instrumental in verifying the
%% properties enforced by the architectures through an encoding into open
%% pNets.
%% We also compare the practical performances of the two algorithms and
%% show the benefit gained using the smart approach.}

>>>>

Figure~\ref{schema:ArchFailure:BIP} shows this modified version of the
Failure Monitor architecture used in~\cite{CubETH-case-study}.\footnote{%
%
  \Simon{%
  The original CubETH case study \cite{CubETH-case-study} focused on
  the possibility of assembling a model of a complex software system
  in a systematic way by applying architectures to discharge
  individual system requirements.  Here, we are more interested in the
  properties of the Failure Monitor architecture by itself.  For that
  reason, we provide a modified version, in particular decomposing the
  coordinator into a Controller and a Timer, altering them to allow
  more flexibility in the acceptable behaviours.  In other words, this
  modification enforces similar properties, while
  discarding less acceptable behaviours.
  }
%
} The architecture consists of 2 \emph{coordinator components}
($\NameCtrl$ and $\NameTimer$, for \emph{Control} and \emph{Timer},
respectively), 5 \emph{dangling ports} ($\PortFail$, $\PortResume$,
$\PortTick$, $\PortReset$ and $\PortAsk$) and 5 connectors.
%
%% Contrary to standard BIP models, architectures comprise one or several
%% \emph{operand} components, whereof only the set of \emph{ports} is
%% given.  Here, the operand component is $\NameOprnd$ and its interface
%% consists of the ports $\PortResume$ and $\PortFail$.
%% The two \emph{coordinator} components\mdash $\NameCtrl$ and
%% $\NameTimer$\mdash are standard
The  \emph{behaviour} of BIP components %% insofar as they also
%% have their 
is specified by finite automata extended with
local data variables.  Transitions of these automata are labelled with
the ports of the corresponding components, Boolean guards and update
functions on local variables.  %% For instance, the loop transition
%% $t_1 \goesto{\PortTick, [t < Max], t:=t+1} t_1$
%% in the $\NameTimer$ component is labelled by the port $\PortTick$. It
%% can be fired only when the current value of the local variable $t$ is
%% lower than $Max$.  Upon firing, this transition increments the value
%% of $t$ by $1$.  When omitted, the default guard (\resp update
%% function) is the constant predicate $\true$ (\resp the $\mathit{skip}$
%% operator).  The constant $\mathrm{Max}$ above, is a parameter of the
%% architecture.
Although this is not visible in the figure, below we will assume that
the dangling ports $\PortFail$ and $\PortResume$ belong to the operand
component, whereas the dangling ports $\PortTick$, $\PortAsk$ and
$\PortReset$ represent actions of the environment.

@@

We validate the approach on \Simon{a larger example}, %\st{two larger examples}, 
based on a BIP architecture taken from an earlier nanosatellite case
study~\cite{CubETH-case-study}.  \Simon{This example shows that
open-automata-based semantics can be instrumental in verifying the
properties enforced by the architectures through an encoding into open
pNets.
In order to evaluate the potential
%
of our approach to scale to larger systems, we also apply the proposed
techniques to a BIP architecture obtained by combining two instances
of the Failure Monitor architecture.  Our results show that, even
though, the performance of the smart algorithm degrades more than
linearly, execution times remain reasonably small.  It should be noted
that the aim of our approach, particularly in the case of BIP
architectures, is to allow symbolic verification of design patterns to
show that they do impose corresponding properties. Safety of the system is obtained compositionally, whereas liveness is verified by checking \emph{pair-wise} non-interference \cite{AttieBBJS16-architectures-faoc}. In both cases, the main characteristic of the approach is compositionality, so we can compute the behavior and analyse the properties on \emph{small open systems}, and use these properties independently when composing bigger systems.
%
}
%% \st{We also show how such architectures
%% can be composed, and how our approach scales.
%% This example shows that
%% open-automata-based semantics can be instrumental in verifying the
%% properties enforced by the architectures through an encoding into open
%% pNets.
%% We also compare the practical performances of the two algorithms and
%% show the benefit gained using the smart approach.}
